Natural Language Interfaces to Structured Data

### 2012-08-27 tr-nlisurvey/src/nli-survey.tex
commit

\usepackage{multirow}
\usepackage{url}
\usepackage{color}

@@

\subsubsection{Query languages}
Relational databases are generally associated with the SQL query language (and
temporal databases are associated to an extension of SQL called temporal SQL).
Multidimensional databases (e.g., data warehouses) are associated with MDX,
which is a SQL-like query languages dedicated for handling measures, dimensions
and hierarchies (key concepts of multidimensional models).
Other data stuctures are associated with their own query language. For instance,
hierarchical lists (or specification lists) is not associated with a language,
but to a similar structure, which is a template where empty slots correspond to
the expected items of the resultset. XML databases are associated to XPath, and
semantic databases (e.g., RDF) are usually queried with languages derived from
SPARQL which looks similar to SQL.
Recently, new query languages have appeared (``no-SQL''). 
\TODO{Develop\ldots}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OVERVIEW SECTION                    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Taxonomy of main approaches}
\label{sec:overview}
% \begin{figure*}
% \centering
% \includegraphics[width=12cm]{images/query-language}
% \caption{Blabla}
% \label{fig:query-languages}
% \end{figure*}

\textbf{Classic translation approach}
The classic translation approach consists in going from (1) to (4).
To reach step (3), semantic rules are needed to find out what database elements
should be associated to question phrases.

\textbf{Iterative approach}
The iterative approach does not go directly from (1) to (3), but goes back to
(1) several times before reaching step (3). 
Those iterations correspond to question reformulations, that involve
user-feedback to interpret semantically the question in terms of a database
query.

@@

\begin{table*}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
%\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{\bf Approach} & \multirow{2}{*}{\bf System} &
\multicolumn{6}{c|}{\bf Dimensions}\\\cline{3-8}
& & {\bf Query} & {\bf Data} & {\bf Answers} & {\bf Portability} & {\bf
Linguistic coverage} & {\bf Error feedback}\\\hline\hline 
\multirow{2}{*}{Domain-dependent semantic parsing} & {\sc
Baseball}~\cite{Green:1961:BAQ:1460690.1460714} & & & & & & \\\cline{2-8}
& {\sc Lunar}~\cite{Woods:1973:PNL:1499586.1499695} & & & & & & \\\hline
\multirow{5}{*}{Complex question translation} & {\sc
Chat-80}~\cite{Warren:1982:EEA:972942.972944} & & & & & & \\\cline{2-8}
& {\sc Irus}~\cite{Bates:1983:IRU:511793.511804} & & & & & & \\\cline{2-8}
& {\sc Qwerty}~\cite{Nelken:2000:QTD:992730.992808} & & & & & & \\\cline{2-8}
& {\sc
Precise}~\cite{Popescu:2004:MNL:1220355.1220376,Popescu:2003:TTN:604045.604070}
& & & & & & \\\cline{2-8}
& {\sc Panto}~\cite{Wang:2007:PPN:1419662.1419706} & & & & & & \\\hline
\multirow{5}{*}{Feedback-driven approaches} &
{\sc Masque/SQL}~\cite{Androutsopoulos93masque} & & & & & & \\\cline{2-8}
& {\sc Team}~\cite{Grosz:1987:TED:25672.25674} & & & & & & \\\cline{2-8}
& {\sc NaLIX} \& {\sc DaNaLIX}~\cite{Li:2005:NIN:1066157.1066281} & & & & & &
\\\cline{2-8}
& {\sc C-Phrase}~\cite{Minock:2010:CSB:1715942.1716190} & & & & & &
\\\cline{2-8}
& {\sc
Orakel}~\cite{Cimiano:2007:PNL:1216295.1216330} & & & & & & \\\hline
\multirow{3}{*}{Learning-based approaches} & Zettlemoyer
et al.~\cite{DBLP:conf/uai/ZettlemoyerC05} & & & & & & \\\cline{2-8}
& Miller et al.~\cite{Miller:1996:FSA:981863.981871} & & & & & & \\\cline{2-8}
& {\sc Wolfie}~\cite{Thompson:2003:AWM:1622420.1622421} & & & & & &
\\\hline
\multirow{2}{*}{Schema-unaware approaches} & {\sc
PowerAqua}~\cite{DBLP:conf/esws/LopezMU06} & & & & & & \\\cline{2-8}
 & {\sc DeepQA}~\cite{DBLP:journals/aim/FerrucciBCFGKLMNPSW10} & & & & & &
 \\\hline
\end{tabular}
\caption{Overview of major NL interfaces to structured data}
\label{tab:overview}
\end{table*}

\subsection{Domain-dependent semantic parsing}
Early systems belong to this class of approaches.
The knowledge necessary to translate NL questions to database queries
is encoded in lexicon.
The systems allows users to use a subset of English (controlled language) to
query databases. 
The lexicon, however, is a huge linguistic resource which defines how each word
must be translated into database elements, and what semantic rules to trigger to
get, in the end, the desired database query.

>>>>

% multirow
\usepackage{multirow}

% url
\usepackage{url}

% allow colors for inline comments
\usepackage{color}

% draw pictures with tikz package
\usepackage{tikz}
\usetikzlibrary{arrows, petri, topaths}
\usepackage{tkz-berge}

\usepackage{rotating}

@@

\subsubsection{Query languages}
Relational databases are generally associated with the SQL query language (and
temporal databases are associated to an extension of SQL called temporal SQL).
Multidimensional databases (e.g., data warehouses) are associated with MDX,
which is a SQL-like query languages dedicated for handling measures, dimensions
and hierarchies (key concepts of multidimensional models).
Other data stuctures are associated with their own query language. For instance,
hierarchical lists (or specification lists) is not associated with a language,
but to a similar structure, which is a template where empty slots correspond to
the expected items of the resultset. XML databases are associated to XPath, and
semantic databases (e.g., RDF) are usually queried with languages derived from
SPARQL which looks similar to SQL.
Recently, new query languages have appeared (``no-SQL'').

@@

\begin{figure*}
\centering
\begin{tikzpicture}[transform shape]
%\tikzset{VertexStyle/.append style = {minimum size = 3pt}}
\node at (0,0) [draw,circle,inner sep=2pt](1){1};
\node at (3,0) [draw,circle,inner sep=2pt](2){2};
\node at (7,0) [draw,circle,inner sep=2pt](3){3};
\node at (10,0) [draw,circle,inner sep=2pt](4){4};
\node at (14,0) [draw,circle,inner sep=2pt](5){5};
\node at (0,0) [draw=none,fill=none,label={[label distance=3pt]0:{NL query}}]
{}; 
\node at (3,0) [draw=none,fill=none,label={[label
distance=3pt]0:{intermediate query}}] {}; 
\node at (7,0) [draw=none,fill=none,label={[label
distance=3pt]0:{query graph}}] {};
\node at (10,0) [draw=none,fill=none,label={[label
distance=3pt]0:{logic query graph}}] {};
\node at (14,0) [draw=none,fill=none,label={[label
distance=3pt]0:{database query}}] {};
\draw[->](1) edge[bend left] (2);
\draw[->](2) edge[bend left] (3);
\draw[->](3) edge[bend left] (4);
\draw[->](4) edge[bend left] (5);
\draw[<->](1) edge[bend right] (3);
\draw[->](4) edge[bend left] (3);
\end{tikzpicture}
\caption{Big picture of NLI systems}
\label{fig:big-picture}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OVERVIEW SECTION                    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Taxonomy of main approaches}
\label{sec:overview}

Figure~\ref{fig:big-picture} depicts the big picture of translating NL queries
to database queries.

\textbf{Classic translation approach}
The classic translation approach consists in going from
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}
to \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}} through
 \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}},
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}
and 
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}} (step
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}} being optional).

To reach step (3), semantic rules are needed to find out what database elements
should be associated to question phrases.

\textbf{Iterative approach}
The iterative approach does not go directly from
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}} to
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}, but goes back to
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} (and/or to
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}) several times before
reaching step \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}.
Those iterations correspond to question reformulations, that involve
user-feedback to interpret semantically the question in terms of a database
query.

@@

\begin{table}
\centering
\rotatebox{90}{
\begin{minipage}{\textheight}
\begin{tabular}{|c|l|c|c|c|c|c|c|}\hline
%\begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
\multirow{2}{*}{\bf Approach} & \multirow{2}{*}{\bf System}
&
\multicolumn{6}{c|}{\bf Dimensions}\\\cline{3-8}
& & {\bf Query} & {\bf Data} & {\bf Answers} & {\bf Portability} & {\bf
Linguistic coverage} & {\bf Error feedback}\\\hline\hline 
\multirow{2}{*}{Domain-dependent semantic parsing} & {\sc
Baseball}~\cite{Green:1961:BAQ:1460690.1460714} & & & & & & \\\cline{2-8}
& {\sc Lunar}~\cite{Woods:1973:PNL:1499586.1499695} & & & & & & \\\hline
\multirow{5}{*}{Complex question translation} & {\sc
Chat-80}~\cite{Warren:1982:EEA:972942.972944} & & & & & & \\\cline{2-8}
& {\sc Irus}~\cite{Bates:1983:IRU:511793.511804} & & & & & & \\\cline{2-8}
& {\sc Qwerty}~\cite{Nelken:2000:QTD:992730.992808} & & & & & & \\\cline{2-8}
& {\sc
Precise}~\cite{Popescu:2004:MNL:1220355.1220376,Popescu:2003:TTN:604045.604070}
& & & & & & \\\cline{2-8}
& {\sc Panto}~\cite{Wang:2007:PPN:1419662.1419706} & & & & & & \\\hline
\multirow{5}{*}{Feedback-driven approaches} &
{\sc Masque/SQL}~\cite{Androutsopoulos93masque} & & & & & & \\\cline{2-8}
& {\sc Team}~\cite{Grosz:1987:TED:25672.25674} & & & & & & \\\cline{2-8}
& {\sc NaLIX} \& {\sc DaNaLIX}~\cite{Li:2005:NIN:1066157.1066281} & & & & & &
\\\cline{2-8}
& {\sc C-Phrase}~\cite{Minock:2010:CSB:1715942.1716190} & & & & & &
\\\cline{2-8}
& {\sc
Orakel}~\cite{Cimiano:2007:PNL:1216295.1216330} & & & & & & \\\hline
\multirow{3}{*}{Learning-based approaches} & Zettlemoyer
et al.~\cite{DBLP:conf/uai/ZettlemoyerC05} & & & & & & \\\cline{2-8}
& Miller et al.~\cite{Miller:1996:FSA:981863.981871} & & & & & & \\\cline{2-8}
& {\sc Wolfie}~\cite{Thompson:2003:AWM:1622420.1622421} & & & & & &
\\\hline
\multirow{2}{*}{Schema-unaware approaches} & {\sc
PowerAqua}~\cite{DBLP:conf/esws/LopezMU06} & & & & & & \\\cline{2-8}
 & {\sc DeepQA}~\cite{DBLP:journals/aim/FerrucciBCFGKLMNPSW10} & & & & & &
 \\\hline
\end{tabular}
\end{minipage}
}
\caption{Overview of major NL interfaces to structured data}
\label{tab:overview}
\end{table}

\subsection{Domain-dependent semantic parsing}
Early systems belong to this class of approaches.
The knowledge necessary to translate NL questions to database queries
is encoded in lexicon.
The systems allow users to use a subset of English (controlled language) to
query databases. 
The lexicon, however, is a huge linguistic resource which defines how each word
must be translated into database elements, and what semantic rules to trigger to
get, in the end, the desired database query.
